FINE SCALE REGISTRATION OF SIDEWALKS AND OTHER RIBBON-LIKE FEATURES FROM AERIAL IMAGERY

### 2018-07-27 Chapters/1.Introduction/Introduction.tex
Fixing format and images

\section{Motivations}
Aerial imagery is an important resource for applications that include urban planning and map-making \cite{5523977}. Ribbon-like features such as sidewalks, street networks, canals, or biking paths are especially important because they provide essential connections between sites to support transportation of people and materials. Any damage or obstacle appeared on these ribbon-like features may cause problems or hurt individual people. High-resolution aerial imagery has become a way to capture the appearance of these structures in order to look for changes such as new construction, degradation or damage, or obstructions. Geometrically, walking paths are thin features that can be described by a trajectory with varying thickness\cite{10.1007/11744078_9}. The scope of this thesis is automatic refinement of ribbon-like features: given an approximate trajectory and range of potential thicknesses, we automate the process of estimating the thicknesses and trajectory. One example of a downstream application would be an automatic method for precise localization and assessment of the quality of these pathways from geometric referenced imagery. Which could enable accurate walking instructions for people including those with disabilities that prevent them from moving around or seeing obstacles\cite{ZOU2012227}.

\begin{figure}
    \centering
    \includegraphics[width=0.57\textwidth]{Figures/Goal.png}
    \caption[Objective Demonstration]{A demonstration on our approach. The first row shows the given input and the second row shows the output, with edges marked in red. We output the edges information as a .geojson file which allow us to edit and change with the given input file.}
    \label{fig:goal}
\end{figure}

Public domain or crowd-sourced repositories of spatial data such as \ac{OSM} \cite{OpenStreetMap} represent a very complete and frequently updated source of information which includes street vector data, and to a much lesser extent, the locations of foot-ways. However, the data is acquired from a variety of sources that may not be registered with enough precision to accurately place thin features such as walking paths. For thin structures such as sidewalks that are visible in \ac{VHR} imagery, even small discrepancies between the digitized pathways and the imagery can be distracting when used to visualize an overland map. As new \ac{VHR} aerial imagery becomes available thin features such as walking paths are discernible with higher precision, but ortho-rectification processes may not precisely align the imagery with previously digitized features. In addition, poorly registered walking paths could confound any algorithm that might hope to automatically locate features on the path including obstacles, damage, overgrowth, or stairways that could impact the walk-ability. In addition, several authors\cite{femiani2009interval, femiani2007road} have worked on segmenting streets and sidewalks or other linear features directly from aerial imagery; however sidewalks are much thinner than streets and are often partially occluded. More importantly, these methods are mainly concerned with the connectivity of the extracted pathways and not with identifying the precise boundaries. Figure \ref{fig:goal} shows the input and desired output with our approach, we marked the boundaries as red. Figure \ref{fig:fw_ov} demonstrate the step by step detail diagram for our approach, from the input map (a), along with the initial trajectory(b), we can generate the ribbon-image with the given mid-curve(c). We generate the initial estimation for each pixels(d) with \ac{GMM} along the ribbon-image, then apply our approach with \ac{DP} solution(e). Our approach can be used to refine or recover the thickness of street networks or other networks of ribbon-like features. Several automatic approaches model segmentation as a \ac{CRF} in order to ensure that the resulting output configurations are plausible \cite{ActiveContou09, Rother2004-ou, Achanta:149300}.

\begin{figure}[H]
    \centering
    \includegraphics[width=0.75\textwidth]{Figures/diagram6.png}
    \caption[Framework Overview 1]{Our 6 step approach adopted to predict precise boundaries for ribbon-like features. (a) Input Map, (b) Initial Trajectory, (c) Ribbon Image Generation, (d) Density Estimation for Pixels, (e) Apply \ac{DP} Solution, (f) Reshape Sidewalk.}
    \label{fig:fw_ov}
\end{figure}

@@

\begin{figure}[H]
    \centering
    \includegraphics[width=0.8\textwidth]{Figures/challenge.png}
    \caption[Challenges Demonstration]{A detail demonstrations on challenges that we are facing in our approach. From left to right, top to bottom, it shows that sidewalk are under the shadow, camouflage with adjacent materials and blocked by obstacles.}
    \label{fig:challenge_demo}
\end{figure}

>>>>

\section{Motivation}
Aerial imagery is an important resource for applications that include urban planning and map-making~\cite{5523977}. 
Ribbon-like features such as sidewalks, street networks, canals, or biking paths are especially important because they provide essential connections between sites to support transportation of people and materials. 
Any damage or obstacle that appears on these ribbon-like features may interfere with the navigability of the walking path and could even cause harm to people.
High-resolution aerial imagery has become a way to capture the appearance of these structures in order to look for changes such as new construction, degradation, damage, or obstructions. 
Geometrically, walking paths are thin features that can be described by a trajectory with varying thickness~\cite{10.1007/11744078_9}. 
The scope of this thesis is automatic refinement of ribbon-like features: given an approximate trajectory and range of potential thicknesses, we automate the process of estimating the thicknesses and offset of the trajectory. 
One example of a downstream application is to enable accurate walking instructions for people including those with disabilities that prevent them from moving around or seeing obstacles~\cite{ZOU2012227}.

\begin{figure}
    \centering
    \includegraphics[width=0.85\textwidth]{Figures/Goal.png}
    \caption[Objective Demonstration]{A demonstration  on input data and output result. By given the input map and initial trajectory (left), we produce the precious boundaries(right) with our \ac{DP} approach. The trajectory mark in black and the boundaries mark in red. We output the edges information as a .geojson file which allow us to edit and change with the given input file.}
    \label{fig:goal}
\end{figure}

Public domain or crowd-sourced repositories of spatial data such as \ac{OSM}~\cite{OpenStreetMap} represent a very complete and frequently updated source of information which includes street vector data, and to a much lesser extent, the locations of foot-ways.
However, the data is acquired from a variety of sources that may not be registered with enough precision to accurately place thin features such as walking paths. For thin structures such as sidewalks that are visible in \ac{VHR} imagery, even small discrepancies between the digitized pathways and the imagery can be distracting when used to visualize an overlaid map.
As new \ac{VHR} aerial imagery becomes available thin features such as walking paths are discernible with higher precision, but ortho-rectification processes may not precisely align the imagery with previously digitized features.

% \todo[]{show result with initial trajectory is off}

In addition, poorly registered walking paths could confound any algorithm that might hope to automatically locate features on the path including obstacles, damage, overgrowth, or stairways that could impact the walk-ability. 
Several authors\cite{femiani2009interval, femiani2007road} have worked on segmenting streets and sidewalks or other linear features directly from aerial imagery; however sidewalks are much thinner than streets and are often partially occluded. 
More importantly, these methods are mainly concerned with the connectivity of the extracted pathways and not with identifying the precise boundaries. 
Figure \ref{fig:goal} shows the input map with the initial trajectory (left) and result from our approach (right), we marked initial trajectory in black and boundaries in red.
Figure \ref{fig:fw_ov} demonstrate the step by step detail diagram for our approach, from the input map (a), along with the initial trajectory(b), we can generate the ribbon-image with the given mid-curve(c). 
We generate the initial estimation for each pixels(d) with initial density along the ribbon-image, then apply our approach with \ac{DP} solution(e). 
Our approach can be used to refine or recover the thickness of street networks or other networks of ribbon-like features. Several automatic approaches model segmentation as a \ac{CRF} in order to ensure that the resulting output configurations are plausible \cite{ActiveContou09, Rother2004-ou, Achanta:149300}.

\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{Figures/diagram.png}
    \caption[Framework Demonstration]{
    Our approach adopted to predict precise boundaries for ribbon-like features. Where black line indicated a sidewalk's geometric information, and red lines shown our result for the sidewalk boundaries. From (a) to (f), the 6 step approach introduce input map (a), initial trajectory (b), ribbon image generation (c), density Estimation for pixels(d), the result from our \ac{DP} solution(e), the result after reshape sidewalk(f).}
    \label{fig:fw_ov}
\end{figure}

@@

\begin{figure}[H]
    \centering
    \includegraphics[width=0.75\textwidth]{Figures/Challenge.png}
    \caption[Challenges Demonstration]{A detail demonstrations on challenges that we are facing in our approach. From left to right, top to bottom, it shows that sidewalk are under the shadow, camouflage with adjacent materials and blocked by obstacles.}
    \label{fig:challenge_demo}
\end{figure}

### 2018-07-27 Chapters/2.Theory/Background.tex
Fixing format and images

\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{Figures/arz_7_resolution_compare.png}
    \caption[Resolution Comparison]{Comparison between different resolutions. From left to right, where (a) has 300 dpi, (b) has 200 dpi and (c) has 100 dpi. With the number of dpi decrease, it harder to capture the boundary of the edge detail of sidewalk.}
    \label{fig:vhr_compare}
\end{figure}

\ac{VHR} stands for Very High Resolution, it's advantageous for buildings extraction, road detection or image segmentation\cite{FREIRE20141, 10.1007/BFb0015525, 10.1007/978-3-642-15567-3_16}. These aerial imageries contain more information and detail due to the higher resolution, which gives us the ability to process more complex analysis on the imagery with same size. \figref{fig:vhr_compare} demonstrates the difference between the same area with different resolution. From (a) to (c), the resolution changes from 300 \ac{dpi} to 100 \ac{dpi}, with higher resolution, it's easier to find the precious boundaries of an object or identify it. For building extraction, with the technique to combine 2-D edge information, photometric and chromatic attributes with 3-D information \cite{10.1007/BFb0015525}, higher resolution provide more information when dealing with complex houses or extract complex suburban roofs. For thin ribbon-features such as sidewalks in our approach, when to visualize an overland map, without high resolution, the imagery can be distracting, and the output may effect by tiny discrepancies between digitized  pathways\cite{10.1007/978-3-642-15567-3_16}.

\section{\ac{OSM}}
\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{Figures/oxford_path_data_osm.png}
    \caption[\ac{OSM} Walking Path]{A demonstration on extract the path data from \ac{OSM}. This is a time-saving process by avoiding hand mark all the sidewalk location with given data set. The figure shows in the Oxford area in Ohio, US.}
    \cite{OpenStreetMap}
    \label{fig:osm_oxford_path}
\end{figure}
\ac{OSM} stands for Open Street Map, which founded by Steve Coast in 2004 \cite{lasPiñas, OpenStreetMap}. \ac{OSM} is a collaborative project that allows users to generate a free editable map of the world \cite{4653466}. Rather than the map itself, \ac{OSM} generate the data with map information that across most of the world as primary output. It also allows users to develop the map data with portable satellite navigation devices. With all these convenience features, \ac{OSM} is helpful for map analysis and feature segmentation \cite{10.1007/11744078_9}. \figref{fig:osm_oxford_path} shows that \ac{OSM} supports features like footpaths, roads, buildings, and others, with the ability to extract the \ac{GIS} data for any individual feature. It's a time-saving process when dealing with large data set, by given the ability to locate the \ac{GIS} data for each feature that we need to analyze. It may lead to the issue of dealing with missing or invalid data when we rely on the data set for auto-detection, or mismatch data with imagery when structures changes along time.

\section{\ac{CRF}}
\ac{CRF} stands for Conditional Random Field, which is a class of statistical modeling method that often applied in pattern recognition and machine learning \cite{MAL-013}. As a type of discriminative undirected probabilistic graphical model, \ac{CRF} is commonly used for segmenting, labeling or parsing of sequential data, such as in computer vision, structure segmentation and object recognition \cite{RuizSarmiento2015UPGMppA, 1315232, lafferty2001conditional}. For road segmentation, several approaches are able to locate the precise boundaries by treating it as a \ac{CRF} problem \cite{ActiveContou09, Achanta:149300}.

@@

\ac{GMM} stands for Gaussian Mixture Model \cite{sridharan2014gaussian}.It is a probabilistic model for representing normally distributed subpopulations within an overall population. \ac{GMM} often can be performed when the data points are mixtures of a Gaussian distribution. It can be used in common computer vision tasks such as background and foreground subtraction \cite{1333992}. For road segmentation, we can treat the problem as subtract foreground (road feature) from the complex background (non-road feature). \figref{fig:gmm_sample_1} demonstrate the comparison between foreground and background \ac{GMM} on sample walking path show in grayscale. For foreground subtraction, which shown in row 1 and row 3, we treat sidewalk texture as foreground and apply the \ac{GMM} algorithm to subtract it from the background (non-sidewalk texture). Background subtraction are shown in row 2 and row 4, we did the opposite and treat non-sidewalk part as foreground.

\begin{figure}[H]
    \centering
    \includegraphics[width=0.8\textwidth]{Figures/gmm_sample_1.png}
    \caption[\ac{GMM} Background Subtraction]{A demonstration of apply \ac{GMM} algorithm on sample inputs, show with gray scale. From top to bottom, row 1 shows foreground \ac{GMM} on sample path, row 2 show background \ac{GMM} on the same sample path as row 1. Same with row 3 and 4, with different sample path. For foreground \ac{GMM}, lighter color represent more likely to be sidewalk and darker color represent more likely to be non-sidewalk. Background \ac{GMM} shows as the opposite.}
    \label{fig:gmm_sample_1}
\end{figure}

\section{\ac{DP}}

\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{Figures/QVSdv.png}
    \caption[Demonstration on Fibonacci Sequence]{Representing the calculation of Fibonacci Sequence ($7^{th}$) with binary tree. Without applying dynamic programming, it need's certain exponential rate to calculate $n^{th}$ Fibonacci number. With the increase of $n$, the number of calculation increase significantly.}
    % \cite{stack_overflow}
    \label{fig:fbs}
\end{figure}

\ac{DP} stands for Dynamic Programming. It's both a mathematical optimization method and a computer programming method \cite{bertsekas2005dynamic}. In 1950s \ac{DP} was invented by Richard Bellman\cite{bellman2013dynamic}. The main concept of \ac{DP} is to simplify a complicated problem by breaking it down into simpler sub-problems in a recursive manner \cite{howard1966dynamic}. In computer science, if a problem can break into serial sub-problems and each sub-problem can be recursively solved by finding it's optimal solution, then we know that \ac{DP} are applicable.

Dynamic Programming has a variety of applications since it can significantly reduce the run-time \cite{bertsekas1995neuro}. We use Fibonacci Sequence to demonstrate the \ac{DP} process\cite{horadam1961generalized}. Figure \ref{fig:fbs} shows that to calculate the $n^{th}$ Fibonacci number, it require a certain exponential rate of calculation to finish the process. The time complexity can be reduced since it's re-calculating the previous Fibonacci number repeatedly. By applying \ac{DP} algorithm, we can save the previous calculated Fibonacci number and reused it when needed. Figure \ref{fig:fbs_dp} indicate the same process after applying dynamic programming, we marked unnecessary steps in red X. By trade in linear space to save previous calculation, we can reduce the wrong time from exponential to linear.

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{Figures/QVSdv_DP.png}
    \caption[Demonstration on Fibonacci Sequence with \ac{DP}]{Representing the calculation of Fibonacci Sequence ($7^{th}$) with binary tree. With dynamic programming, after saving previous calculation, it's only require linear time steps to calculate $n^{th}$ Fibonacci number. We marked unnecessary steps in red X to show it's can be skipped.
    }
    % \cite{stack_overflow}
    \label{fig:fbs_dp}
\end{figure}

>>>>

\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{Figures/arz_7_res_compare.png}
    \caption[Resolution Comparison]{Comparison between different resolutions. From left to right, where (a) has half meter scales, (b) has 5 meter scales and (c) has 10 meter scales. With the increase of meter scale, it harder to capture the boundary of the edges detail of sidewalks.}
    \label{fig:vhr_compare}
\end{figure}
% \todo{change dpi to 5 meter scale. make it worse.}

\ac{VHR} stands for Very High Resolution, it has advantageous for buildings extraction, road detection or image segmentation~\cite{FREIRE20141, 10.1007/BFb0015525, 10.1007/978-3-642-15567-3_16}. 
These aerial imageries contain more information and detail due to the higher resolution, which give us the ability to develop more complex analysis on the imagery with same size.
\figref{fig:vhr_compare} demonstrates the difference between different resolution within the same area. From (a) to (c), the resolution changes from half meter scale to 10 meter scale. 
With higher resolution, it's easier to find the precious boundaries of an object or identify it. 
For building extraction, higher resolution provides more information when dealing with complex houses or extract complex suburban roofs~\cite{10.1007/BFb0015525}. 
For thin ribbon-features segmentation such as sidewalks in our approach, to visualize an overlaid map without high resolution, the imagery can be distracting, and the output may effect by tiny discrepancies between digitized pathways~\cite{10.1007/978-3-642-15567-3_16}.

\section{\ac{OSM}}
\begin{figure}[H]
    \centering
    \includegraphics[width=0.85\textwidth]{Figures/oxford_path_data_osm.png}
    \caption[\ac{OSM} Walking Path]{A demonstration of actual walking path data from \ac{OSM}. The figure shows in the Oxford area in Ohio, US. Row 2 specifies the zoom in version of error that we need to facing. The sidewalk misses in the path data (left), the path data marks offset compare to the actual sidewalk(right).}
    \cite{OpenStreetMap}
    \label{fig:osm_oxford_path}
\end{figure}

\ac{OSM} stands for Open Street Map, which was founded by Steve Coast in 2004~\cite{lasPiñas, OpenStreetMap}. 
\ac{OSM} is a collaborative project that allows users to generate a free editable map of the world~\cite{4653466}.
Rather than the map itself, \ac{OSM} generates the data with the map information that across most of the world as primary output. 
It also allows users to develop the map data with portable satellite navigation devices. 
With all these convenience features, \ac{OSM} is helpful for the map analysis and features segmentation~\cite{10.1007/11744078_9}. 
\figref{fig:osm_oxford_path} shows that \ac{OSM} supports features like footpaths, roads, buildings, and others, with the ability to extract the \ac{GIS} data for any individual feature. 
It may also leads to the issue of dealing with missing or invalid data when we rely on the data set for auto-detection, or mismatch data with imagery when structures changes along with time.

\section{\ac{CRF}}
\ac{CRF} stands for Conditional Random Field, which is a class of statistical modeling method that often applied in pattern recognition and machine learning~\cite{MAL-013}. 
As a type of discriminative undirected probabilistic graphical model, \ac{CRF} is commonly used for segmenting, labeling or parsing of sequential data, such as in computer vision, structure segmentation and object recognition~\cite{RuizSarmiento2015UPGMppA, 1315232, lafferty2001conditional}. 
For road segmentation, several approaches are able to locate the precise boundaries by treating it as a \ac{CRF} problem~\cite{ActiveContou09, Achanta:149300}.

@@

\ac{GMM} stands for Gaussian Mixture Models~\cite{sridharan2014gaussian}.
It is a probabilistic model for representing normally distributed subpopulations within an overall population.
\ac{GMM} often can be performed when the data points are mixtures of a Gaussian distribution. 
It can be used in common computer vision tasks such as background and foreground subtraction~\cite{1333992}. 
For road segmentation, we can treat the problem as subtract foreground (road feature) from the complex background (non-road feature). 
\figref{fig:gmm_sample_1} demonstrate the comparison between foreground and background \ac{GMM} on sample walking path show in gray-scale. 
For foreground subtraction, which shown in row 1 and row 3, we treat sidewalk texture as foreground and apply the \ac{GMM} algorithm to subtract it from the background (non-sidewalk texture). 
Background subtraction are shown in row 2 and row 4, we did the opposite and treat non-sidewalk part as foreground.

\begin{figure}[H]
    \centering
    \includegraphics[width=0.85\textwidth]{Figures/GMM_needed.png}
    \caption[Density Background Subtraction]{A demonstration of apply \ac{GMM} algorithm on sample inputs, show with gray scale. From top to bottom, row 1 shows foreground density on sample path, row 2 show background density on the same sample path as row 1. Same with row 3 and 4, with different sample path. For foreground density, lighter color represent more likely to be sidewalk and darker color represent more likely to be non-sidewalk. Background density shows as the opposite.}
    \label{fig:gmm_sample_1}
\end{figure}

\section{\ac{DP}}
\begin{figure}
    \centering
    \includegraphics[width=0.9\textwidth]{Figures/longest_common_sequence.png}
    \caption[Demonstration on Longest Common Sequence]{Dynamic Programming demonstration with longest common sequence. Similar to our approach, to find the longest common sequence, it first needs to fill the table with \ac{DP} and then backtrack to find the longest path. It used in intersection detection with GPS trace~\cite{Xie2017DetectingRI}.
    }
    \label{fig:lcs_dp}
\end{figure}

\ac{DP} stands for Dynamic Programming. It is both a mathematical optimization method and a computer programming method \cite{bertsekas2005dynamic}. 
In 1950s \ac{DP} was invented by Richard Bellman\cite{bellman2013dynamic}. 
The main concept of \ac{DP} is to simplify a complicated problem by breaking it down into simpler sub-problems in a recursive manner \cite{howard1966dynamic}. 
In computer science, if a problem can break into serial sub-problems and each sub-problem can be recursively solved by finding its optimal solution, then we know that \ac{DP} are applicable.

\figref{fig:lcs_dp} gives a example of a famous dynamic programming problem, Longest Common Sequence. To achieve the result of finding the longest common sequence, it first needs to fill the table with \ac{DP} and then backtrack to find the longest path. One method gives approach with using longest common sequence for intersection detection from GPS traces~\cite{Xie2017DetectingRI}.

Dynamic Programming has a variety of applications since it can significantly reduce the run-time \cite{bertsekas1995neuro}. 
We use Fibonacci Sequence to demonstrate the \ac{DP} process\cite{horadam1961generalized}. 
Figure \ref{fig:fbs}, row 1 shows that to calculate the $n^{th}$ Fibonacci number, it require a certain exponential rate of calculation to finish the process.
The time complexity can be reduced since it is re-calculating the previous Fibonacci number repeatedly. 
By applying \ac{DP} algorithm, we can save the previous calculated Fibonacci number and reused it when needed.
Figure \ref{fig:fbs}, row 2 indicate the same process after applying dynamic programming, we marked unnecessary steps in red X. 
By trade in linear space to save previous calculation, we can reduce the wrong time from exponential to linear.

\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{Figures/treewithdp.png}
    \caption[Demonstration on Fibonacci Sequence]{Representing the calculation of Fibonacci Sequence ($7^{th}$) with binary tree. Row 1 shows without applying dynamic programming, it need's certain exponential rate to calculate $n^{th}$ Fibonacci number. With the increase of $n$, the number of calculation increase significantly. Row 2 shows with dynamic programming, after saving previous calculation, it's only require linear time steps to calculate $n^{th}$ Fibonacci number. We marked unnecessary steps in red X to show it's can be skipped. }
    % \cite{stack_overflow}
    \label{fig:fbs}
\end{figure}

% \begin{figure}
%     \centering
%     \includegraphics[width=\textwidth]{Figures/QVSdv_DP.png}
%     \caption[Demonstration on Fibonacci Sequence with \ac{DP}]{Representing the calculation of Fibonacci Sequence ($7^{th}$) with binary tree. With dynamic programming, after saving previous calculation, it's only require linear time steps to calculate $n^{th}$ Fibonacci number. We marked unnecessary steps in red X to show it's can be skipped.
%     }
%     % \cite{stack_overflow}
%     \label{fig:fbs_dp}
% \end{figure}

### 2018-07-27 Chapters/2.Theory/Theory.tex
Fixing format and images

A variety of approaches are able to locate precise object boundaries, although few directly address the problem of sidewalk segmentation in "blobby" or compact regions. Several tools for segmentation appear for segmenting ribbon-like features in \ac{VHR}, we list Grabcut \cite{Rother2004-ou}, Slic\cite{Achanta:149300}, and Active Contours \cite{Kass88snakes:active} as examples.
Results for these methods on a simple sidewalk are shown in \figref{fig:Method_comparison} which shows qualitative results of each method. We observe that grabcut had decent performance on this example, but it is still unable to identify edges in shadow or correctly determine the width of the sidewalk region in the vicinity of a driveway, which seems to be made of the same material as the sidewalk. In addition we observe that all three approaches recognize the gutter as part of the sidewalk because it has the similar texture.  
\begin{figure}[H]
    \centering
    \includegraphics[width=0.55\textwidth]{Figures/compare.png}
    \caption[Method comparison with Grabcut, Active Contours, and Slic]{
        % Method comparison with Grabcut, Active contour and Slic
        A comparison between Grabcut(left), SLIC (center) and Active Contours (right) 
        on an example of a sidewalk from 6-inch aerial imagery; 
        \aclp{TP} are rendered in green, 
        \aclp{FN} are rendered in blue, and 
        \aclp{FP} are rendered in red.
        The SLIC image contains all super-pixels that intersect the initial trajectory of a walking path; the Active Contours and GrabCut methods are initialized with the SLIC results.
    }
    \label{fig:Method_comparison}
\end{figure}

@@

Grabcut is a powerful image segmentation tool that combines both texture(color) and edge(contrast) information \cite{Rother2004-ou}. It was designed to efficiently extract a foreground object from a complex background. According to figure \ref{fig:Method_comparison} and figure \ref{fig:grabcut}, the overall performance on grabcut applied to a ribbon-like feature is fairly accurate, but it's hard to identify objects above the sidewalk such as shadows, gutters or foliage. Grabcut seems to identify the occluded regions as negative, we suspect that is because of the lack of width control.

@@

Simple linear iterative clustering (SLIC) works with super-pixels, which segment a large image into different irregularly shaped regions of homogeneous appearance or texture \cite{Achanta:149300}. It uses the compact evenly distributed homogeneous regions of an image as a pre-processing step, so that image segmentation can be accomplished by grouping super-pixels which snap to natural boundaries in an image. Alternatively, super-pixels can initiate a region growing process\cite{Borovec2017-fz}. As shown in figure \ref{fig:slic}, slic will identify a non-sidewalk feature as positive if there is a shadow covering the sidewalk, which may lead to over mark the boundary (\aclp{FP}). This becomes a problem since all super-pixels that we decide to use are based on a line in between both edges of sidewalk which we marked as ground truth. On the one hand, if the super-pixel above lay on the line, we got an error, since we marked part of the feature that not the sidewalk on the sidewalk, which gives us \aclp{FP}. On the other hand, if that super-pixel does not lay on the line, we missed a part that's clear is the sidewalk. Thus, finding a way to separate such a feature become important. Trying to decrease the size of the width of each super-pixel was one of the ways to fix it but may cause an incompleteness since each super-pixel is smaller than the width of the sidewalk. So the idea of width control become important.

@@

A very successful region growing process is snakes, or Active Contours, which starts with an approximate contour and evolves the boundary to optimize an energy function that favors likely contour shapes. Compared with the other two segmentation tools, active contour may not be the best solution for identifying ribbon-like feature since it is not the best tool to separate blurred edges. Al-Diri \cite{ActiveContou09} proposed using active contour to identify each edge separately, in the mean time keep maintaining width consistency. It is a good approach but it is not ideal to apply on our problem, since their ribbon-like feature (vessels) does not requiring predict edges under obstacles, such as trees or cars that above the sidewalk.

Both \ac{SLIC} and active contours start from an initial shape to find precise boundaries by solving a \ac{CRF} problem. In addition, graph cuts are capable of global \ac{CRF} problem. In addition, graph cut are capable of global \ac{CRF} optimization in some situations, and iterative graph cut are used by the GrabCut algorithm to refine an initial segmentation \cite{Rother2004-ou}. However in figure \ref{fig:Method_comparison} and figure \ref{fig:ac}, we show the segmentation results of these methods, and note that they struggle to constrain their results to ribbon-like shapes with smooth thickness and mid-line. These color or boundary smoothness methods fail to capture an important quality of ribbon-like features, which is the thickness at one point depends not only its adjacent points but also its topologically distant points along the contour.

@@

Another image segmentation approach that has seen extraordinary success is the use of \ac{CNNs} and transposed convolution \cite{Badrinarayanan2017-il, Noh2015-ni, Shelhamer2017-rf, Ronneberger2015-sv}. Unlike the proposed approach, these have millions of tunable parameters and need significant pre-annotated data and offline training to build a model. In a recent exploration of learning to segment street networks from online maps \cite{Kaiser2017-np}, built a fully convolutional network with 137 million parameters. A similar approach with another deep architecture called U-net\cite{Ronneberger2015-sv} was applied to the street network segmentation problem \cite{Zhang2017-gi}. These approaches claim to be robust to occlusion and ambiguous regions, but they benefit from a large amount of carefully annotated data that includes road width \cite{Mnih2013-dp}. This type of pre-annotated data with aligned aerial imagery is much harder to obtain for walking paths. The proposed approach has only handful of tunable parameters with clear interpretations, so that they can be interactively set and used without offline training.

>>>>

A variety of approaches are able to locate precise object boundaries, although few directly address the problem of sidewalk segmentation in "blobby" or compact regions. 
Several tools for segmentation appear for segmenting ribbon-like features in \ac{VHR}, we list Grabcut ~\cite{Rother2004-ou}, Slic~\cite{Achanta:149300}, and Active Contours~\cite{Kass88snakes:active} as examples.
Results for these methods on a simple sidewalk are shown in \figref{fig:Method_comparison} which shows qualitative results of each method. 
We observe that grabcut had decent performance on this example, but it is still unable to identify edges in shadow or correctly determine the width of the sidewalk region in the vicinity of a driveway, which seems to be made of the same material as the sidewalk. 
In addition we observe that all three approaches recognize the gutter as part of the sidewalk because it has the camouflage texture.

\begin{figure}[H]
    \centering
    \includegraphics[width=0.55\textwidth]{Figures/compare.png}
    \caption[Method comparison with Grabcut, Active Contours, and Slic]{
        % Method comparison with Grabcut, Active contour and Slic
        A comparison between Grabcut(left), SLIC (center) and Active Contours (right) 
        on an example of a sidewalk from 6-inch aerial imagery; 
        \aclp{TP} are rendered in green, 
        \aclp{FN} are rendered in blue, and 
        \aclp{FP} are rendered in red.
        The SLIC image contains all super-pixels that intersect the initial trajectory of a walking path; the Active Contours and GrabCut methods are initialized with the SLIC results.
    }
    \label{fig:Method_comparison}
\end{figure}

@@

Grabcut is a powerful image segmentation tool that combines both texture(color) and edge(contrast) information~\cite{Rother2004-ou}. 
It was designed to efficiently extract a foreground object from a complex background. 
According to figure \ref{fig:Method_comparison} and figure \ref{fig:grabcut}, the overall performance on grabcut applied to a ribbon-like feature is fairly accurate, but it's hard to identify objects above the sidewalk such as shadows, gutters or foliage. 
Grabcut seems to identify the occluded regions as negative, we suspect that is because of the lack of width control.

@@

Simple linear iterative clustering (SLIC) works with super-pixels, which segment a large image into different irregularly shaped regions of homogeneous appearance or texture~\cite{Achanta:149300}. 
It uses the compact evenly distributed homogeneous regions of an image as a pre-processing step, so that image segmentation can be accomplished by grouping super-pixels which snap to natural boundaries in an image. 
Alternatively, super-pixels can initiate a region growing process~\cite{Borovec2017-fz}. 
As shown in figure \ref{fig:slic}, slic will identify a non-sidewalk feature as positive if there is a shadow covering the sidewalk, which may lead to over mark the boundary (\aclp{FP}).
This becomes a problem since all super-pixels that we decide to use are based on a line in between both edges of sidewalk which we marked as ground truth. 
On the one hand, if the super-pixel above lay on the line, we got an error, since we marked part of the feature that not the sidewalk on the sidewalk, which gives us \aclp{FP}. 
On the other hand, if that super-pixel does not lay on the line, we missed a part that's clear is the sidewalk. Thus, finding a way to separate such a feature become important. 
Trying to decrease the size of the width of each super-pixel was one of the ways to fix it but may cause an incompleteness since each super-pixel is smaller than the width of the sidewalk.
So the idea of width control become necessary.

@@

A very successful region growing process is snakes, or Active Contours, which starts with an approximate contour and evolves the boundary to optimize an energy function that favors likely contour shapes. 
Compared with the other two segmentation tools, active contour may not be the best solution for identifying ribbon-like feature since it is not the best tool to separate blurred edges. 
Al-Diri~\cite{ActiveContou09} proposed using active contour to identify each edge separately, in the mean time keep maintaining width consistency. 
It is a good approach but it is not ideal to apply on our problem, since their ribbon-like feature (vessels) does not requiring predict edges under obstacles, such as trees or cars that above the sidewalk.

Both \ac{SLIC} and active contours start from an initial shape to find precise boundaries by solving a \ac{CRF} problem. 
In addition, graph cut are capable of global \ac{CRF} optimization in some situations, and iterative graph cut are used by the GrabCut algorithm to refine an initial segmentation~\cite{Rother2004-ou}. However in figure \ref{fig:Method_comparison} and figure \ref{fig:ac}, we show the segmentation results of these methods, and note that they struggle to constrain their results to ribbon-like shapes with smooth thickness and mid-line.
These color or boundary smoothness methods fail to capture an important quality of ribbon-like features, which is the thickness at one point depends not only its adjacent points but also its topologically distant points along the contour.

@@

Another image segmentation approach that has seen extraordinary success is the use of Convolutional Neural Networks (CNNs) and transposed convolution~\cite{Badrinarayanan2017-il, Noh2015-ni, Shelhamer2017-rf, Ronneberger2015-sv}. 
Unlike the proposed approach, these have millions of tunable parameters and need significant pre-annotated data and offline training to build a model. 
In a recent exploration of learning to segment street networks from online maps~\cite{Kaiser2017-np}, built a fully convolutional network with 137 million parameters.
A similar approach with another deep architecture called U-net~\cite{Ronneberger2015-sv} was applied to the street network segmentation problem~\cite{Zhang2017-gi}.
These approaches claim to be robust to occlusion and ambiguous regions, but they benefit from a large amount of carefully annotated data that includes road width~\cite{Mnih2013-dp}. 
This type of pre-annotated data with aligned aerial imagery is much harder to obtain for walking paths. 
The proposed approach has only handful of tunable parameters with clear interpretations, so that they can be interactively set and used without offline training.
